Who - The Manual

### 2010-04-03 doc/manual/manual.tex
an option for versions
version 0.3

@@

A type definition of a polymorphic type looks as follows:
\begin{alltt}
  type array ['a||]
\end{alltt}
Notice the two bars at the end to indicate empty region and effect parameters.
Also note that despite the name, {\tt array} is {\em not} a mutable type.

We can declare functions over this type in the following way:
\begin{alltt}
  logic create ['a||] :  int -> 'a -> 'a array
  logic get ['a||] : int -> 'a array -> 'a
  logic set ['a||]: int -> 'a -> 'a array -> 'a array 
  logic len ['a||] :  'a array -> int
\end{alltt}
Again, notice that these are pure functions which act on {\em functional}
arrays, for example {\tt set} returns a new, modified array.

Now, we can declare properties of these functions in form of axioms:
\begin{alltt}
  axiom update_length ['a||] : 
    forall (t : 'a array) (i : int) (z : 'a).
    0 <= i \verb|/\| i < len t -> len t = len (set i z t)
  axiom get_set_neq ['a||] : 
    forall (t : 'a array ) (i : int) (j : int) (z : 'a).
      (0 <= i \verb|/\| i < len t) -> 
      (0 <= j \verb|/\| j < len t) -> 
        i <> j -> get i (set j z t) = get i t
  axiom get_set_eq ['a||] : 
    forall (t : 'a array) (i : int) (z : 'a).
      0 <= i \verb|/\| i < len t -> get i (set i z t) = z
\end{alltt}

Now, we can also {\em define} functions over those functional arrays:
\begin{alltt}
  let zeroset ['a||] (v : 'a) (a : 'a array) = set 0 v a
\end{alltt}
Here, we have defined a {\em pure} function {\tt zeroset} of type {\tt ['a||].
int -> 'a array -> 'a array}. No effects take place, and nothing has to be
proved.

@@

If we want to define a {\em calculatory function} which does the same, here we
go:
\begin{alltt}
  let zerosetc ['a||] (v : 'a) (a : 'a array) = 
    \{ \}
    set 0 v a
    \{ \}
\end{alltt}
Notice the additional curly brackets at the beginning and the end of the
function body. The function {\tt zerosetc} has the type {\tt ['a||]. int -> 'a
array ->{} 'a array}. We see that the last (and only the last) arrow has an
effect annotation. In this case, this effect annotation is empty; no effect
takes place, because {\tt set} returns a new function.

Now, if we want to state anything about the return value of {\tt zerosetc}, we
need to write a {\em postcondition}. For example, we would like to state that
the returned array contains zero at position $i$:
\begin{alltt}
  let zerosetc ['a||] (v : 'a) (a : 'a array) = 
    \{ \}
    set 0 v a
    \{ r : get 0 r = v \}
\end{alltt}
We use the syntax {\tt r : } to speak about the return value; {\tt r} is a
name chosen by the user. This definition produces a proof obligation. In this
case, this proof obligation is actually unprovable, because the array {\tt a}
might be empty, in which case the axioms do not say anything of {\tt set 0 v a}.
We have to give a {\em precondition}:
\begin{alltt}
  let zerosetc ['a||] (v : 'a) (a : 'a array) = 
    \{ 0 < len a \}
    set 0 v a
    \{ r : get 0 r = v \}
\end{alltt}
And now we can prove this function.

@@

>>>>

\chapter{The command line tool}

@@

A type definition of a polymorphic type looks as follows:
\begin{alltt}
  type array ['a]
\end{alltt}
Also note that despite the name, {\tt array} is {\em not} a mutable type.

We can declare functions over this type in the following way:
\begin{alltt}
  logic create ['a] :  int -> 'a -> 'a array
  logic get ['a] : int -> 'a array -> 'a
  logic set ['a]: int -> 'a -> 'a array -> 'a array
  logic len ['a] :  'a array -> int
\end{alltt}
Again, notice that these are pure functions which act on {\em functional}
arrays, for example {\tt set} returns a new, modified array.

Now, we can declare properties of these functions in form of axioms:
\begin{alltt}
  axiom update_length ['a] :
    forall (t : 'a array) (i : int) (z : 'a).
    0 <= i \verb|/\| i < len t -> len t = len (set i z t)
  axiom get_set_neq ['a] :
    forall (t : 'a array ) (i : int) (j : int) (z : 'a).
      (0 <= i \verb|/\| i < len t) ->
      (0 <= j \verb|/\| j < len t) ->
        i <> j -> get i (set j z t) = get i t
  axiom get_set_eq ['a] :
    forall (t : 'a array) (i : int) (z : 'a).
      0 <= i \verb|/\| i < len t -> get i (set i z t) = z
\end{alltt}

Now, we can also {\em define} functions over those functional arrays:
\begin{alltt}
  let zeroset ['a] (v : 'a) (a : 'a array) = set 0 v a
\end{alltt}
Here, we have defined a {\em pure} function {\tt zeroset} of type {\tt ['a].
int -> 'a array -> 'a array}. No effects take place, and nothing has to be
proved.

@@

If we want to define a {\em calculatory function} which does the same, here we
go:
\begin{alltt}
  let zerosetc ['a||] (v : 'a) (a : 'a array) =
    \{ \}
    set 0 v a
    \{ \}
\end{alltt}
Notice the additional curly brackets at the beginning and the end of the
function body. The function {\tt zerosetc} has the type {\tt ['a]. int -> 'a
array ->{} 'a array}. We see that the last (and only the last) arrow has an
effect annotation. In this case, this effect annotation is empty; no effect
takes place, because {\tt set} returns a new function.

Now, if we want to state anything about the return value of {\tt zerosetc}, we
need to write a {\em postcondition}. For example, we would like to state that
the returned array contains zero at position $i$:
\begin{alltt}
  let zerosetc ['a] (v : 'a) (a : 'a array) =
    \{ \}
    set 0 v a
    \{ r : get 0 r = v \}
\end{alltt}
We use the syntax {\tt r : } to speak about the return value; {\tt r} is a
name chosen by the user. This definition produces a proof obligation. In this
case, this proof obligation is actually unprovable, because the array {\tt a}
might be empty, in which case the axioms do not say anything of {\tt set 0 v a}.
We have to give a {\em precondition}:
\begin{alltt}
  let zerosetc ['a] (v : 'a) (a : 'a array) =
    \{ 0 < len a \}
    set 0 v a
    \{ r : get 0 r = v \}
\end{alltt}
And now we can prove this function.

@@

%TODO read/write effects
